---
title: "In-Class Data Exploration 2"
output: html_notebook
authoe: "David Murillo"
date: "September 27, 2022"
---

#Load packages
```{r}
require(here)
require(psych)
```

#Load datasets
```{r}
dat_bird <- read.csv(here("data", "bird.sta.csv"))
dat_habitat <-  read.csv(here("data", "hab.sta.csv"))

```

#Bird habitat data exploration
```{r}
colnames(dat_habitat)
```
#Bird count data
Use the hist() function to explore the distributions of counts of birds at the various study sites.

```{r}
hist(dat_bird$CBCH)
```
I used the arguments xlab = "Number of birds counted" and breaks = 0:7 - 0.5 with hist() to produce the following:

```{r}
hist(dat_bird$CBCH, xlab =  "Number of birds counted", breaks = 0:7 - 0.5)
```
#Bird Count Histograms
Create some histograms of bird counts. Experiment with the breaks argument.

You may want to try to re-create my histogram above as a first step.

If your breaks argument does not span the range of count values you will get an error. You may need to try a higher upper limit.
You could also use the max() function to determine the highest number of counts of your bird species column.

```{r}
hist(dat_bird$WIWR, breaks = 0:8, ylab = "Number of birds counted",
     main = "Histogram of Troglodytes hiemalis")
```


## Instructions
As a group you should work through the following:

#Import the datasets, and take a look at the metadata.
#Explore the habitat data using pair plots.

```{r}
pairs(dat_habitat[, c("elev", "slope", "ba.con", "ba.tot")])
```


Explore the distributions of counts of bird species using histograms.

```{r}
hist(dat_bird$WIWR, breaks = 0:8, ylab = "Number of birds counted",
     main = "Histogram of Troglodytes hiemalis")

hist(dat_bird$WIWA, xlab = "Number of Birds Counted", main = "Wilson's Warbler Abundance", breaks = 0:7 - 0.5, col = "gold2") 
```


