---
title: 'Lab 10: ANOVA (by hand)'
author: "David Murillo"
date: "November 2, 2022"
output:
  html_document:
    theme: cosmo
---

```{r, echo=FALSE, include=FALSE, warning=FALSE}
library(here)

rope <- read.csv(here("data", "rope.csv"))

# Factorial Experiment

#For the purposes of this exercise, we will create a simple one-way ANOVA with:

# predictor: rope type
# response: percent rope cut
```


```{r, echo=FALSE, include=FALSE, warning=FALSE}
# Fill in the code as you go.

rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)

n_obs = sum(!is.na(rope$rope.type))
n_groups = 6

ss_tot = sum( (mean(rope$p.cut) - rope$p.cut)^2)
df_tot = n_obs - 1

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN =  function(x){
    sum( (mean(x) - x)^2)
  } )

ss_within = sum(agg_sq_resids$x)

df_within = n_obs - n_groups

ss_among = ss_tot - ss_within
df_among = df_tot - df_within

ms_within = ss_within / (n_obs - n_groups)
ms_among  = ss_among / (n_groups - 1)

f_ratio = ms_among / ms_within

f_pval = pf(f_ratio, df1 = df_among, df2 = df_within, lower.tail = FALSE)
```

```{r, echo=FALSE, include=FALSE, warning=FALSE}
# Anova

fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)

anova_fit_1 = anova(fit_1)
str(anova_fit_1)

anova_fit_1$"Sum Sq"
```

```{r, echo=FALSE, include=FALSE, warning=FALSE}

# Self test

# number comparison tolerance
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```

# Question {.tabset .tabset-pills}

## ANOVA by hand 

- **Q1 (8 pts.): Submit the code you used to build your ANOVA by hand. Make sure you use the code template so that you use the same variable names as those which I’ll use for the grading.**

*Answer: *

```{r}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)

n_obs = sum(!is.na(rope$rope.type))
n_groups = 6

ss_tot = sum( (mean(rope$p.cut) - rope$p.cut)^2)
df_tot = n_obs - 1

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN =  function(x){
    sum( (mean(x) - x)^2)
  } )

ss_within = sum(agg_sq_resids$x)

df_within = n_obs - n_groups

ss_among = ss_tot - ss_within
df_among = df_tot - df_within

ms_within = ss_within / (n_obs - n_groups)
ms_among  = ss_among / (n_groups - 1)

f_ratio = ms_among / ms_within

f_pval = pf(f_ratio, df1 = df_among, df2 = df_within, lower.tail = FALSE)
```


## Self-check

```{r}
# number comparison tolerance
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```


## Model Assumptions

- **Q2 (1 pt.): Examine the conditional boxplot in the Partitioning Variance: Within-Group section of the walkthrough. Based on the figure, do you think there are equal variances among the groups?**

```{r, echo=FALSE, include=FALSE, warning=FALSE}

boxplot(p.cut ~ rope.type, data= rope, main = "Conditional boxplot of percentage of cut versus rope type", ylab = "Percentage of cut", xlab = "Rope type")

```


*Answer: Looks that the varience is different bewteen rope types, the Blaze, SB and, VEL type look that had more variance that BS and XTC, and PI has the lowest variance*

- **Q3 (1 pt.): Conduct a Bartlett test to assess the homogeneity of variances of the percent cut among the rope type groups. Report the p-value.** 

```{r, echo=FALSE, include=FALSE, warning=FALSE}
bartlett.test(p.cut ~ rope.type, data= rope)
```


*Answer: p-value = 0.00143*

- **Q4 (2 pts.): Given your graphical assessment (question 2) and the Bartlett test, do you think an ANOVA-type analysis is appropriate on the raw data? Explain why or why not.**

*Answer: No, I do not think that one way ANOVA-type analysis was appropriate with the rope data, because the assumption of homocedasticy is does not met. *


## Model coefficients and group means 

- **Q5 (1 pt.): Which rope type is the base case?**

*Answer: The BLAZE rope type*

- **Q6 (1 pt.): What is the mean percent cut of the base case rope? Show your calculation using value(s) from the model coefficient table.**

*Answer: 0.36714*

- **Q7 (1 pt.): What is the mean percent cut rope type XTC? Show your calculation using value(s) from the model coefficient table.**

*Answer: 0.2655*

```{r}
0.36714 - 0.10164
```


## Q 8 pt (R-code)

*Answer: *

```{R }

rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)

n_obs = sum(!is.na(rope$rope.type))
n_groups = 6

ss_tot = sum( (mean(rope$p.cut) - rope$p.cut)^2)
df_tot = n_obs - 1

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN =  function(x){
    sum( (mean(x) - x)^2)
  } )

ss_within = sum(agg_sq_resids$x)

df_within = n_obs - n_groups

ss_among = ss_tot - ss_within
df_among = df_tot - df_within

ms_within = ss_within / (n_obs - n_groups)
ms_among  = ss_among / (n_groups - 1)

f_ratio = ms_among / ms_within

f_pval = pf(f_ratio, df1 = df_among, df2 = df_within, lower.tail = FALSE)

# number comparison tolerance
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```
