---
title: "In-Class Probability 3: Calculations In R"
author: "David Murillo"
date: "October 10, 2022"
output:
  html_document:
    theme: cerulean

---

# Question {.tabset .tabset-pills}

## Binomial Probabilities

- **Q1 (1 pt.): If I wanted to use a binomial distribution to model my six forest plots, what values should I use for the two parameters of a binomial distribution?**

*Answer:*
```{r}
n = 6
p = 4/6

dbinom(1, size= n ,pro = p)
```


- **Q2 (1 pt.): Use dbinom to calculate the probability of observing birds in exactly four of the six patches. Include your R-code in your answer.**

*Answer:*

```{r}
n = 6
p = 4/6

dbinom(4, size= n ,pro = p)
```

- **Q3 (1 pt.): Now, suppose I did a survey and observed no birds in my plots. Use dbinom to calculate the probability of observing no presences.**

*Answer:*
```{r}
n = 6
p = 4/6

dbinom(0, size= n ,pro = p)
```


## Binomial cumulative probabilities: p-functions

- **Q4 (1 pt.): Back to the binomial scenario (bird presence/absence in 6 forest plots).**

*Answer:*
```{r}
n = 6
p= 4/6
pbinom(q = 4, size = n , p= p)


```


Now use pbinom to calculate the probability of observing four or fewer presences in the 6 plots. Show your R code.

- **Q5 (1 pt.): Now use pbinom and the law of total probability to calculate the probability of observing four or more presences in the 6 plots. Show your R code.**

*Answer:*
```{r}
n = 6
p= 4/6
1 - pbinom(q = 3, size = n , p= p)


```
```


## Normal Probabilities

- **Q6 (1 pt.): Are you more likely to observe a value of 1.0 or 2.0?**

*Answer: 1*

- **Q7 (1 pt.): What is the probability of observing a value of 1.0 or less? Show the R code you used to find your answer.**

*Answer:*

```{r}
pnorm(1, mean= 0, sd= 1)

```


- **Q8 (1 pt.): What is the probability of observing a value between 1.0 and 2.0? Show the R code you used to find your answer.**

*Answer:*

0.135

```{r}

(pnorm(2, mean = 0, sd= 1))  - (pnorm(1, mean = 0, sd = 1))


```


## Normal plots

- **Q9 (2 pts.): Show the complete R-code you used to create your plot. Make sure you include all the code to recreate your plot in a fresh R session.**

*Answer:*

```{r}
n = 1000

x = seq(from = -10, to = 10, length.out = n)

y = dnorm(x, mean = 0, sd = 1)
y2 = dnorm(x, mean = 0, sd = 2)
y3 = dnorm(x, mean = -2, sd = 1)

# plot!
plot(y ~ x, type = "l", ylab = "Probability Density")
plot(y2 ~ x, type = "l", add= TRUE)
plot(y3 ~ x, type = "l", add= TRUE)

y2 = pnorm(x, mean = 0, sd = 1)

# plot!
plot(y2 ~ x, type = "l", ylab = "Probability Density")



```


- **Q10 (1 pt.): Include a figure of your plot.**

*Answer:*

## Binomial plots

- **Q11 (2 pts.): Show the complete R-code you used to create your plot. Make sure you include all the code to recreate your plot in a fresh R session.**

*Answer:*

- **Q12 (1 pt.): Include a figure of your plot.**

*Answer:*

