---
title: "Lab 03: Data Exploration and Deterministic Functions"
output: html_notebook
author: "David Murillo"
date: "September/21/2022"
---

# Load packages
```{r}
library(psych)
library(here)
```

# Load datasets
```{r}
dat_bird <- read.csv("data/bird.sta.csv")
dat_habitat <- read.csv("data/hab.sta.csv")

```

# Merge data
```{r}
dat_all <- merge(dat_habitat, dat_bird)

plot(ba.tot ~ elev, data = dat_all)
```

1. Use the pair plot function from psych to create a pair plot of the three terrain variables and basal area from the lecture questions.
```{r}
pairs.panels(dat_all[,c("ba.con", "ba.hard","ba.snag", "ba.tot")])
```


2. Choose two bird species and create plots of presence/absence (on the y-axis) and basal area (on the x axes).

```{r}
BGWA <- as.numeric(dat_all$BGWA) > 0

SWTH <- as.numeric(dat_all$SWTH) > 0
```

Visually inspect the plots and fit logistic curves using the parameterization functions provided above.

```{r}
# Function to calculate the logistic parameter a given the slope and midpoint
get_logistic_param_a = function(slope, midpoint)
{
  b = slope / 4
  return (-midpoint * (slope / 4))
}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
{
  return (slope / 4)
}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{
  val = exp(a + b * x)
  return(val / (1 + val))
}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))
}
```

```{r}
plot(dat_all$elev, BGWA)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.1), add = TRUE)

plot(dat_all$elev, SWTH)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = 0.1), add = TRUE)
```


 How many total number of Gray Jays were observed in all of the sampling sites.
 
```{r}
sum(dat_all$GRJA)
```
 Show the R code you used to perform the calculation.
 
 Calculate the total number of sampling sites in which Gray Jays were observed.
 
```{r}
 GRJA_sta <- sum(as.numeric(dat_all$GRJA > 0))
GRJA_sta
 
```
 
