---
title: 'Lab 06: Introduction to Inference'
author: "David Murillo"
date: "Octuber 12, 2022"
output:
  html_document:
    theme: cosmo
---

```{r, echo=FALSE, warning=FALSE, include=FALSE}
#R Function for the SE

sse_mean = function(x){
  sd(x, na.rm = TRUE )/sqrt( length(x))
}

require(palmerpenguins)

sse_mean(penguins$bill_depth_mm)

```
```{r, echo=FALSE, warning=FALSE, include=FALSE}
# 2-species data

dat_pen = subset(penguins, species != "Gentoo")

boxplot(
  flipper_length_mm ~ species, data = dat_pen,
  ylab = "Flipper length (mm)")

#droplevels()

dat_pen = droplevels(subset(penguins, species != "Gentoo"))

par(mfrow = c(1, 2))
  boxplot(
    flipper_length_mm ~ species, data = penguins,
    ylab = "Flipper length (mm)")

```


```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Resampling with replacement
#What if we randomly shuffled the flipper length data and made another boxplot?

# for reproducibility
set.seed(123)

flipper_shuffled = sample(
  penguins$flipper_length_mm, replace = TRUE)

{
  par(mfrow = c(1, 2))
  boxplot(
    flipper_length_mm ~ species, data = penguins,
    ylab = "Flipper length (mm)",
    main = "Original Data")
  boxplot(
    flipper_shuffled ~ penguins$species,
    ylab = "Flipper length (mm)",
    main = "MonteCarlo Resampled Data",
    xlab = "species")
}
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Bootstrap Resampling and Alternative Hypotheses

penguins2 = penguins[sample(1:nrow(penguins), replace = T), ]

{
  par(mfrow = c(1, 2))
  boxplot(
    flipper_length_mm ~ species, data = penguins,
    ylab = "Flipper length (mm)",
    main = "Original Data")
  boxplot(
    flipper_length_mm ~ species, data = penguins2,
    ylab = "Flipper length (mm)",
    main = "Bootstrap Data")
}
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Classical t-test: Adelie and Chinstrap penguins

t.test(dat_pen$flipper_length_mm ~ dat_pen$species)

# Two-sample resampling

# Reset the random number generator state for reproduceablility
set.seed(1)
flipper_shuffled = sample(dat_pen$flipper_length_mm)

boxplot(flipper_shuffled ~ dat_pen$species)
```
```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Using aggregate()

agg_means = aggregate(
  flipper_length_mm ~ species, 
  data = dat_pen, 
  FUN = "mean", 
  na.rm = TRUE)

diff_observed = diff(agg_means[, 2])

agg_means
diff_observed

```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Resampling experiment

# If I ran this function many times, how often would I see a mean difference greater than diff_observed?

```
# Question {.tabset .tabset-pills}

## Sample Standard Error function 

- **Q1 (3 pts.): Show the R code you used to define your sse_mean() function. Include the following line before your function definition:**

*Answer:*

```{r}
rm(list = ls())

sse_mean = function(x){
  sd(x, na.rm = TRUE )/sqrt( length(x))
}

sse_mean(penguins$body_mass_g)
sse_mean(mtcars$mpg)
```

## Two Group Resampling 

- **Q2 (4 pts.): Show the code you used to define your `two_group_resample_diff()` function.**

*Answer:*

```{r}

two_group_resample_diff = function(x, n_1, n_2) {
  
  group1 = mean(sample(x, n_1, replace = TRUE), na.rm = TRUE )
  
  group2 = mean(sample(x, n_2, replace = TRUE), na.rm = TRUE)
  
  return(group1 - group2)
}

```

- **Q3 (2 pts.): Does your function perform Monte Carlo or bootstrap resampling, i.e. does it simulate a null or an alternative hypothesis? You may want to review your answer after you complete the rest of the lab questions.**

*Answer: Perform Monte Carlo resampling, because we get different vaues every time, this simulate a null hypothesis*

## Resampled Flipper Lengths 

Use your `two_group_resample_diff()` function along with a loop to generate 2000 resampled differences of means for flipper length between the two penguin species (Adelie and Chinstrap).

- **Q4 (1 pt.): Create a histogram of the resampled differences of means.**

*Answer:*

```{r, echo= FALSE, warning=FALSE, fig.align='center', fig.width= 6, fig.height=4}

dat_pen <-   penguins

n = 2000

mean_differences = c()

dat_pen <- droplevels(subset(dat_pen, species %in% c( "Adelie", "Chinstrap")))

for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample_diff(dat_pen$flipper_length_mm, 68, 152)
  )
}

hist(mean_differences, main= "Histogram of mean differences",
     xlab = "mean differences")
```

- **Q5 (2 pts.): How many of your resampled differences of means had a magnitude greater than 5.8?**

*Answer: 0 *

```{r}
sum(mean_differences > 5.8)
```

## Resampling and p-values

- **Q6 (2 pts.): Given a p value of less than 1 per 10 million, how many simulations do you think you would have to do to see a difference in mean flipper length equal to or greater than 5.8 mm?**

*Answer: at least 10 million, this because we have a p value of 0.00000006, this mean that if we repeat the analysis with different set of data on 99.9999999% of the time we will see a different less that 5.8 mm, if we want to see a different greater than 5.8 mm we need to redo the analysis at leats 10 million of times.*

## Resampling a different variable

    - Choose another variable of penguin data from dat_pen.

    - Create a boxplot of the data from your chosen column in dat_pen, grouped by species.

```{r, echo=FALSE, include=FALSE, warning=FALSE}
dat_pen = droplevels(subset(penguins, species != "Gentoo"))

par(mfrow = c(1, 2))
  boxplot(
    bill_length_mm ~ species, data = dat_pen,
    ylab = "Bill length (mm)")

```

    - See the walkthrough for info.

    - Use aggregate() to calculate the group means and the difference in means.

    - Write the difference in means into a variable called diff_crit

```{r, echo=FALSE, include=FALSE, warning=FALSE}
agg_means_bill = aggregate(
  bill_length_mm ~ species, 
  data = dat_pen, 
  FUN = "mean", 
  na.rm = TRUE)

diff_crit = diff(agg_means_bill[, 2])

diff_crit
```


    - Conduct a t-test and observe the p-value.

```{r, echo=FALSE, include=FALSE, warning=FALSE}
t.test(dat_pen$bill_length_mm ~ dat_pen$species)

```

    - Conduct a resampling test with 1000 repetitions using your two_group_resample_diff() function.

```{r, echo=FALSE, include=FALSE, warning=FALSE}
dat_pen <- penguins

n = 1000

mean_differences = c()

dat_pen <- droplevels(subset(dat_pen, species %in% c( "Adelie", "Chinstrap")))

for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample_diff(dat_pen$bill_length_mm, 68, 152)
  )
}

hist(mean_differences)

```


- **Q7 (1 pt.): Include a boxplot of your chosen variable in your report.**

*Answer:*
```{r, echo=FALSE, fig.align='center', fig.height=6, fig.width=6}

  boxplot(
    bill_length_mm ~ species, data = dat_pen,
    ylab = "Bill length (mm)",
    main= "Difference on bill length among species")
```


- **Q8 (3 pts.): Report the group means and difference between the means.**

*Answer: 10.04243 *

- **Q9 (4 pts.): Interpret the p-value from the t-test output in plain, non-technical English that a non-statistician would understand.**

*Answer: The p values is 2.2e-16, in normal number is 0.00000000000000022, this value due of the t-test suggests that there is good evidence that the bill length is different between the two species. Our P value is so low that, if we redoing the analysis with different set of data we will need much more than 100 millions of intents to found a different between bill length of penguins.*

*If we use the alpha value of 0.05 to interpret the p value, we could see that if we repeat the analysis with different set of data at least the 95% of the intent we will see a good evidence that the bill length is different between penguin species.*

- **Q10 (2 pts.): How many differences in means were greater than diff_crit?**
  - Remember to treat this as a 2-tailed test.

*Answer: Our diff_crit is 10.04243, and there are 0 diffences in means that were greater that 10.04243*
  
```{r}
sum(mean_differences > diff_crit)
```


- **Q11 (1 pt.): Include a histogram of your simulation results in your report. Make sure it has appropriate title and labels.**

*Answer:*

```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=4}
hist(mean_differences, main= "Histogram of mean differences",
     xlab= "Mean difference")
```





