---
title: "Probability and Frequentist Concepts"
author: "David Murillo"
date: "October 20, 2022"
output:
  html_document:
    theme: cosmo
  html_notebook: default
---

# Questions

### Q1 (2 pts.): What is the probability of observing a count of exactly 3 successes in a binomial distribution with parameters n = 4 and p = 0.75?

*Answer: 0.42*

```{r}
dbinom(x= 3, size= 4, prob = 0.75)

```

### Q2 (2 pts.): What is the probability of observing a count of 3 successes or fewer in a binomial distribution with parameters n = 4 and p = 0.75?

*Answer: 0.683*

```{r}
pbinom(3, size= 4, prob = 0.75, lower.tail = TRUE)
```

### Q3 (2 pts.): What is the probability of observing more than 3 successes in a binomial distribution with parameters n = 5 and p = 0.75?

*Answer: 0.63 *

```{r}
1 - pbinom(3, size= 5, prob = 0.75, lower.tail = TRUE)
```

### Q4 (2 pts.): - What is the probability of observing a value of less than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

*Answer: 0.326 *

```{r}
pnorm(1.1, mean =2, sd = 2, lower.tail = TRUE)
```

### Q5 (2 pts.): - What is the probability of observing a value of greater than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

*Answer: 0.655*

```{r}
1 - pnorm(1.2, mean =2, sd= 2, lower.tail = TRUE)
```

### Q6 (4 pts.): - What is the probability of observing a value between 1.2 and 3.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

*Answer: 0.399 *

```{r}
pnorm(3.2, mean = 2, sd =2, lower.tail = TRUE) - pnorm(1.1, mean = 2, sd= 2, lower.tail = TRUE) 
```

### Q7 (2 pts.): Describe how the shape of the histogram changes as you continue to press the sample button.

*Answer: With the firsts samples the histogram take the shape of the beta distribution, with more sample the histograma get stable, and between more sample the histogram get more stable*

### Q8 (2 pts.): Describe how the shape of the histogram changes as you continue to press the sample button.

*Answer: In this case the histogram does not fit the beta distribution in its entirety, and need much more sample to get stable*

### Q9 (2 pts.): Describe how the shape of the histogram changes as you continue to press the sample button.

*Answer: In this case the histogram does not fit the beta distribution, or is more difficult to look, the majority of the values fall in the center of the histogram*


### Q10 (2 pts.): Why is there such a drastic change in the shape of the sampling distribution when you change the sample size from 1 to 2?

*Answer: Because with one sample the mean values will be equal to the sample value, when we use two sample the mean will be between sample one and sample two values*

### Q11 (2 pts.): What are the two main factors that determine the width of the sampling distribution of the mean?

*Answer: The sample size, and the standard deviation*


### Q12 (2 pts.): How many 3-character words are possible?

There are 25 possible words consisting of a single character in the Library. There are 25×25 =625 possible 2-character words.

*Answer:15625 *

```{r}
25^3
```

### Q13 (2 pts.): How many books would the Library contain if you added one additional position to the book size (i.e. one extra letter on the last page)? Express your answer in terms of B.

*Answer: If we add one additional positions per row, we will get a B value of: 253,724,400*

```{r}
B <- (410 * 40 * 81) * 191
B
```