---
title: 'Code Chunks, Chunk Options, and Tabs'
author: "David Murillo"
date: "Octuber 6, 2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: lumen
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r histogram_penguins, echo = FALSE}
require(palmerpenguins)
hist(penguins$body_mass_g)
```
```{r multi_panel_1, echo= FALSE, fig.height= 10, fig.width= 10}
par(mfrow = c(2, 2))

hist(
penguins$body_mass_g,
main = "Histogram of Body Mass",
xlab = "Body Mass (g)")

hist(
penguins$flipper_length_mm,
main = "Histogram of Flipper Length",
xlab = "Flipper Length (mm)")

boxplot(
body_mass_g ~ species, data = penguins,
main = "Body Mass and Species",
ylab = "Body Mass (g)"
)

boxplot(
flipper_length_mm ~ species, data = penguins,
main = "Flipper Length and Species",
ylab = "Flipper Length (mm)"
)

```
# A New Section With Tabs {.tabset}

## Tab 1

*This is my first tme working with Tabs, Looks that will be funny!*

## Tab 2

Boxplot of flipper length vrs species of pinguins:

```{r tab_graph 1, echo = FALSE}
boxplot(
  flipper_length_mm ~ species, data = penguins,
  main = "Flipper Length and Species",
  ylab = "Flipper Length (mm)"
)

```

## Tab 3

Here's a numbered list:

1. Item 1
1. Item 2
1. Item 3
1. **Item 4** (in bold)

# New 1st Level header (not there are no tabs)


## Subsection 1

This section is not in a tab

## Subsection 2

Still no tabs!

# Tab Section 2 {.tabset .tabset-pills}

This seciton uses a 'pill style' of tabs!

## Introducction to pinguins

The pinguins are birds adapt to swin very weel, they can not fly, but swin like fish.

## Boxplot

Flipeer length vrs species of pinguins:

```{r tab_graph 2, echo = FALSE}
boxplot(
  flipper_length_mm ~ species, data = penguins,
  main = "Flipper Length and Species",
  ylab = "Flipper Length (mm)"
)

```

## The pinguins are

Characteristic:

1. Are birds
1. Can not fly
1. Can swin
1. **Live in very cold place**
