---
title: "In-Class Confidence Intervals"
author: "David Murillo"
date: "October 18, 2022"
output:
  html_document:
    theme: cerulean

---

```{r}
qnorm(c(0.025, 0.975))

qnorm(c(0.05, 0.95))
```

# Q1 90% Standard Normal Critical Values


 - **Q1 (1 pt.): Calculate the critical z-values for a 90% CI of the standard normal distribution (not a 95% interval). Show the R-code you used to perform the calculation.**
 
 
```{r}
qnorm(c(0.05, 0.95))
```
 
 # Q2-4 95% T-distribution critical values
 
 - **Q2 (1 pt.): Consult the help entry for qt() and calculate the critical values for df = 10. Show the R-code you used to perform the calculation. **
 
 
```{r}
alpha = 0.05

lower <- qt(alpha/2, df= 10)
upper <- qt(alpha/2, df= 10)

lower
upper
```
 - **Q3 (2 pts.): How many degrees of freedom are required for the 0.025% lower critical value of a t-distribution to match the 0.025% lower critical z-value (from the standard normal) to within one decimal place? Show the R-code you used to perform the calculation.**
 
```{r}

alpha= 0.05

upper <- qt(alpha/2, df= 450)
lower <- qt(1- alpha/2, df= 450)

upper
lower

```
 
 - **Q4 (1 pt.): How many degrees of freedom are required for the 0.025% lower critical value of a t-distribution to match the 0.025% lower critical z-value (from the standard normal) to within two decimal places? Show the R-code you used to perform the calculation.**
 
```{r}
alpha= 0.05

upper <- qt(alpha/2, df= 450)
lower <- qt(1- alpha/2, df= 450)

upper
lower
```
 
# Q5-6 Building a 95% CI of the mean.

Recall the general procedure for constructing a CI? Check out the last section of slide deck 5 if you need a refresher.

Suppose you know that the sample standard deviation for a group of 50 measurements is 3.14. The mean value is 10.0.

```{r}
ssd = 3.14
n = 50
mean = 10
alpha = 0.05

crit_value <- qt(1- alpha/2, df = n-1 )
crit_value

se <- ssd/sqrt(n)

Upper <- mean + crit_value * se
Lower <- mean - crit_value * se
Upper
Lower
```


- **Q5 (2 pts.): What are the critical t-values you would need to know to construct a 95% CI on the mean?**

```{r}
ssd = 3.14
n = 50
mean = 10
alpha = 0.05

crit_value <- qt(1- alpha/2, df = n-1 )
crit_value
```

 
- **Q6 (3 pts.): Construct the interval. Show the R-code you used to perform the calculation.**

```{r}
se <- ssd/sqrt(n)

Upper <- mean + crit_value * se
Lower <- mean - crit_value * se
Upper
Lower




```

