---
title: 'Lab 07: The Bootstrap'
author: "David Murillo"
date: "Octuber 19, 2022"
output:
  html_document:
    theme: cosmo
---


```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Basic examples

# Create simulated data
dat = matrix(1:49, nrow = 7, ncol = 7)
print(dat)

# Minimum and maximum values in each row

apply(dat, MARGIN = 1, FUN = min)

apply(dat, MARGIN = 2, FUN = max)

# Mean values in each column

apply(dat, MARGIN = 2, FUN= mean)

```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# You’ll need the moths.csv file for the data
library(here)

moths = read.csv(here("data", "moths.csv"))
head(moths)

# For now, let’s focus on just one species, the spiny oakworm, Anisota stigma, abbreviated ‘anst’, and examine the distribution of its abundance with a histogram

hist(moths$anst, xlab= "Abundance", main= "Histogram of Anisota stigma abundance")
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# A Parametric Confindence Interval
# For the purposes of the calculations in this lab, you should use a t-distribution.

# Calculating the Parametric CI

# Choose significance level
alpha = 0.05

# 2: Calculate sample standard error:
n = sum(!is.na(moths$anst))
sse = sd(moths$anst, na.rm = TRUE) / sqrt(n)

# 3: Calculate critical t-values:
t_crit = abs(qt(alpha / 2, df = n - 1))

# 4: Calculate the CI radius:
ci_radius = sse * t_crit

# The CI is the sample mean +/- the radius:
anst_ci = c(
  lower = mean(moths$anst) - ci_radius,
  upper = mean(moths$anst) + ci_radius)

print(round(anst_ci, 4))


```


```{r, echo=FALSE, warning=FALSE, include=FALSE}
# A Simple Bootstrap Confidence Interval
# If we don’t want to use a parametric approach, we can use bootstrapping.

# Calculating the Bootstrap CI

m = 10000

# numeric() creates an vector of length m with all values initialized to zero
result = numeric(m)
head(result)

# Perform the bootstrap

for(i in 1:m)
{
  result[i] = mean(sample(moths$anst, replace=TRUE))
}

# Calculate the quantiles

# The vector result now contains 10,000 bootstrap sample means. We can calculate the mean of the bootstrap means and, more importantly, the 2.5% and 97.5% quantiles of the bootstrap distribution, as follows:

mean(result)

quantile(result,c(0.025,0.975))
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Bootstrap Interval Using boot()
library(boot)

# Custom Mean Function

# To use it with boot() we have to create a modified version of the mean() function that automatically excludes NAs:

boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

# Now we can find the bootstrap for 10000 iterations:

myboot = 
  boot(
    data = moths$anst,
    statistic = boot_mean,
    R = 10000)

print(myboot)

#original is the original mean of the whole sample: mean(moths$anst).
#bias is the difference between the original mean and the mean of the bootstrapped samples.
#std.error is the standard deviation of the simulated values.

# Lastly, we can extract our bootstrap confidence interval as follows:

quantile(
  myboot$t,
  c(0.025, 0.975))
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Rarefaction Curve

# Setting up the bootstrap

moth_dat = moths[,-1]
head(moth_dat)

n = nrow(moth_dat) #number of rows or sample observations

m = 100 #number of bootstrap iterations

moth_result = matrix(
  nrow = m,
  ncol = n)


# The outer loop: runs once for each bootstrap iteration.  index variable is i
for(i in 1:m)
{
  # The inner loop: simulates increasing sampling intensity
  # Sampling intensity ranges from 1 site to the complete count of sites (24)
  # index variable is j
  for(j in 1:n)
  {
    # sample the input data row indices, with replacement
    rows_j = sample(n, size = j, replace=TRUE)

    # Creates a new data matrix from the resampled rows.
    t1 = moth_dat[rows_j, ]

    # Calculates the column sums of the new data matrix.
    t2 = apply(t1, 2, sum)

    # Counts the number of columns in which any moths were observed
    moth_result[i, j] = sum(t2 > 0)
  }
}

head(moth_result)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# First Draft
# In my first draft, I’ll just copy all of the code inside the function and propose argument names based on the function inputs I identified above:

rarefaction_sampler = function(input_dat, n_iterations)
{
  n = nrow(moth_dat) #number of rows or sample observations
  m = 100 #number of bootstrap iterations

  moth_result = matrix(
    nrow = m,
    ncol = n)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:m)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of sites (24)
    # index variable is j
    for(j in 1:n)
    {

      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = moth_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      moth_result[i, j] = sum(t2 > 0)
    }
  }

  return(moth_result)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Second Draft

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    # index variable is j
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)
```


```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Check in a fresh environment
# This clears the current R session's environment
rm(list = ls())

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))
moth_dat = moths[,-1]

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(input_dat = moth_dat, n_iterations =100)
head(rarefact)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
require(here)
# Debugging template

# This clears the current R session's environment
rm(list = ls())

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moths[,-1], 100)
head(rarefact)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Building the Rarefaction Curve

# You are now ready to continue with the rarefaction.

# Run the simulator with 10000 iterations:

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))
rarefact = rarefaction_sampler(moths[,-1], 1000)
head(rarefact)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
#For convenience, let’s bind the objects together and transpose the data frame so that the columns represent the mean, 2.5% and 97.5% quantiles, and the rows represent sampling intensity ranging from 1 to 24.

rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))

upper <- rare_quant[2,] 
lower <-  rare_quant[1,] 
```


```{r, echo=FALSE, warning=FALSE, include=FALSE}
# Plotting the curve

mycol <- rgb(0, 0, 255, max = 255, alpha = 125, names = "blue50")

matplot(
  rare,
  type='l',
  xlab='Number of sampling plots',
  ylab='Species richness',
  main="David's Awesome Rarefaction Curve",
  lwd = c(3,3,3),
  col = c(1,2,3))

legend(
  'bottomright',
  legend=c('mean','2.5%','97.5%'),
  lty=c(1,2,3),col=c(1,2,3), inset=c(.1,.1) )

polygon(c(1:24, rev(1:24)), c(upper, rev(rare_mean)), col = mycol,
        border = NA)

polygon(c(1:24, rev(1:24)), c(lower, rev(rare_mean)), col = mycol,
         border = NA)
```


# Questions {.tabset .tabser-pills}

## Penguin Parametrics CI

- Review the bootstrap and parametric confidence interval materials in the lab walkthrough.

- Calculate a parametric 95% CI for mean bill length (in mm) for the Gentoo penguins in the penguins dataset from package `palmerpenguins` using your SSE function. For this calculation you should use Student’s t-distribution to calculate the critical values

```{r, echo=FALSE, include=FALSE}
library(palmerpenguins)

Gentoo <- subset(penguins, species == "Gentoo")

SSE = function(x){
  sd(x, na.rm = TRUE )/sqrt(sum(!is.na(x)))
}

alpha = 0.05

n = sum(!is.na(Gentoo$bill_length_mm)) 

t_crit = abs(qt(alpha / 2, df = n - 1))

ci_radius <- t_crit * SSE(Gentoo$bill_length_mm)

bill_ci = c(
  lower = mean(Gentoo$bill_length_mm, na.rm = TRUE) - ci_radius,
  upper = mean(Gentoo$bill_length_mm, na.rm = TRUE) + ci_radius)

print(round(bill_ci, 4))
```

- Q1 (1 pt.): What is the sample size, n? Show the code you used for the calculation and remember to check for missing data.

*Answer: n is 123 *

```{r}

n = sum(!is.na(Gentoo$bill_length_mm)) 
```


- Q2 (1 pt.): What is the sample standard deviation? Show the code you used for the calculation.

*Answer: the ssd is 3.081857*

```{r}

sd(Gentoo$bill_length_mm, na.rm = TRUE)
```


- Q3 (2 pts.): What are the critical t-values? Show the R code you used for the calculation.

*Answer: the critical t-value is 1.9796 *

```{r}

t_crit = abs(qt(alpha / 2, df = n - 1))
t_crit
```


- Q4 (1 pt.): What is the sample standard error? Show the R code you used for the calculation.

*Answer: the ssd is 0.2778817*

```{r}

SSE = function(x){
  sd(x, na.rm = TRUE )/sqrt(sum(!is.na(x)))
}

SSE(Gentoo$bill_length_mm)
```


- Q5 (2 pts.): Finally, construct the CI and show the R code you used for the calculation.

*Answer: the lower CI is 46.9548 and the upper is 48.0550 *

```{r}
ci_radius <- t_crit * SSE(Gentoo$bill_length_mm)

bill_ci = c(
  lower = mean(Gentoo$bill_length_mm, na.rm = TRUE) - ci_radius,
  upper = mean(Gentoo$bill_length_mm, na.rm = TRUE) + ci_radius)

print(round(bill_ci, 4))
```


## Penguin Bootstrap CI

    - Review the bootstrap confidence interval materials in the lab walkthrough.
    - Calculate a bootstrap 95% CI for mean bill length (in mm) for the Gentoo penguins in penguins dataset from package palmerpenguins.
    - Use the boot() function from package boot()

```{r, echo=FALSE, include=FALSE}

library(boot)

boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

myboot = 
  boot(
    data = Gentoo$bill_length_mm,
    statistic = boot_mean,
    R = 10000)

print(myboot)

quantile(
  myboot$t,
  c(0.025, 0.975))
```

- Q6 (1 pt.): What is the CI?

*Answer:  the 2.5% is 46.97964, and the 97.5% is 48.05287 *

- Q7 (1 pt.): Show the r code you used to call the boot() function.

*Answer: *

```{r}
library(boot)

boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

myboot = 
  boot(
    data = Gentoo$bill_length_mm,
    statistic = boot_mean,
    R = 10000)

```


- Q8 (2 pts.): Show the r code you used to calculate the upper and lower 2.5% quantiles.

*Answer: *

```{r}

quantile(
  myboot$t,
  c(0.025, 0.975))

```


## Rarefaction Sampler 

    -You’ll be using your rarefaction_sampler() function together with the rare moth abundance data.

    -Review the rarefaction curve materials in the lab walkthrough and complete the the rarefaction_sampler() function code.

    -Be sure to test out your completed function using the debugging template to make sure it works correctly.

    -Calculate a rarefaction curve using 10000 replicate simulations.

    -Create a plot of your moth sampling rarefaction curve using the code in the lab walkthrough as a template. Include a 95% confidence envelope, i.e. upper 97.5 and and lower 2.5%.

- Q9 (5 pts.): Show your completed rarefaction_sampler() function.

*Answer: *

```{r}

rm(list = ls())

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  for(i in 1:n_iterations)
  {
    
    for(j in 1:n_input_rows)
    {
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      t1 = input_dat[rows_j, ]

      t2 = apply(t1, 2, sum)

      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

```


- Q10 (1 pt.): What did you find most difficult about building the function?

*Answer: Was a little complicate to know that the n value have to change for the sample size of x, but was a nice exercise*

- Q11 (4 pts.): Show the code you used to perform the simulations and construct the curve.

**Answer:*

```{r, echo=TRUE, results='hide'}

moths = read.csv(here("data", "moths.csv"))
rarefact = rarefaction_sampler(moths[,-1], 1000)
head(rarefact)

```


- Q12 (4 pts.): Include your rarefaction curve plot in your report. Show the R-code you used to create your plot.

*Answer: *

```{r, fig.height=6, fig.width= 8, fig.align='center'}

rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))

upper <- rare_quant[2,] 
lower <-  rare_quant[1,] 

mycol <- rgb(0, 0, 255, max = 255, alpha = 125, names = "blue50")

matplot(
  rare,
  type='l',
  xlab='Number of sampling plots',
  ylab='Species richness',
  main="Rarefaction Curve",
  lwd = c(3,3,3),
  col = c(1,2,3))

legend(
  'bottomright',
  legend=c('mean','2.5%','97.5%'),
  lty=c(1,2,3),col=c(1,2,3), inset=c(.1,.1) )

polygon(c(1:24, rev(1:24)), c(upper, rev(rare_mean)), col = mycol,
        border = NA)

polygon(c(1:24, rev(1:24)), c(lower, rev(rare_mean)), col = mycol,
         border = NA)
```


- Q13 (2 pts.): About how many sites should you visit if you want to see all of the moth species? Explain your reasoning using your rarefaction curve figure.

*Answer: Between 15 and 20 sites, accord to the rarefaction curve the maximum richness of moth species is reached between the size 15 and 20 *


