---
title: "In-Class Likelihoods"
author: "David Murillo"
date: "October 25, 2022"
output:
  html_document:
    theme: cosmo
  html_notebook: default
---

```{r, echo=FALSE, include=FALSE, warning=FALSE}
#Here’s an example in R:

#Let’s say I’ve been conducting counts of Wilson’s Warblers at study sites in Oregon.
#I’ve been to 2 sites where I counted 2, and 6 birds.

x_observed = c(2, 6)
print(x_observed)

dpois(x = 2, lambda = 4.5)

dpois(x = 6, lambda = 4.5)

dpois(x = 2, lambda = 4.5) * dpois(x = 6, lambda = 4.5)

wiwa_counts = c(2, 6)
dpois(x = wiwa_counts, lambda = 4.5)

prod(dpois(x = wiwa_counts, lambda = 4.5)) # Product of Likelihoods

sum(log(dpois(x = wiwa_counts, lambda = 4.5))) # Sum of Log-Likelihoods


# Likelihood of Many Observations
library(here)

dat_bird = read.csv(here::here("data", "bird.sta.csv"))
dat_habitat = read.csv(here::here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)


```


# Questions

Consider the likelihood calculation example for 2 observations of Wilson’s Warblers using a Poisson distribution model.

- Recall the R-code I used in the walkthrough to find the log-likelihood of using λ=4.5:

As a group, experiment with different values of λ to find a value that maximizes the sum of log-likelihoods of the hypothetical observations.

wiwa_counts = c(2, 6)
dpois(x = wiwa_counts, lambda = 5)

sum(log(dpois(x = wiwa_counts, lambda = 4)))

## **Q1 (1 pt.): What value for λ did you select?**

*Answer: 4 *

## **Q2 (1 pt.): How did you choose a value?**

*Answwer: Because is the mean of 2 and 6*


In the walkthrough, I tried using a Poisson distribution with λ=1.0 to model the set of census counts of Wilson’s Warbler.

You will create models for the census count of Winter Wrens.

- Find the λ value of a Poisson model that makes all of the observed the Winter Wren’s census counts most likely.

```{r}
mean(dat_bird$WIWR)
```


- Plot a histogram of Winter Wren counts (Check the metadata file to find the abbreviation for Winter Wrens). Make sure you choose an appropriate number of bins for the plot.

```{r}
hist(dat_all$WIWR, breaks = 0:7 - .5, 
     main = "Histogram of abundance \n Winter Wrens",
     xlab = "Abundance")
```


- As a group, you’ll be searching for the value of λ that maximizes the sum of log-likelihood. In other words, you’ll find the the model parameter that makes the values of observed wren census counts most likely.

```{r}
sum(log(dpois(x = dat_all$WIWA, lambda = 1.45)))
```

- Hint: you only need to consider 2 significant figures in your value of λ.

## **Q3 (1 pt.): Include your histogram into your report.**

```{r}
png(filename = "WIWR_Histogram.png", width = 6, height = 4, res = 300, units = "cm")

hist(dat_all$WIWR, breaks = 0:7 - .5, 
     main = "Histogram of abundance \n Winter Wrens",
     xlab = "Abundance")

dev.off()

```


**Q4 (1 pt.): What value for λ did you select?**

*Answer:  1.45 *

**Q5 (2 pts.): Show the R code you used to calculate the Poisson log-likelihood for the vector of Winter Wren census counts.**

*Answer: sum(log(dpois(x = dat_all$WIWA, lambda = 1.45))) *


In the walkthrough, I tried using a Poisson distribution with λ=1.0 to model the set of census counts of Wilson’s Warblers.

You will create models for the census count of Winter Wrens.


Suppose you wanted to use a binomial distribution instead of a Poisson distribution.

Find values of the parameters for a binomial distribution that make the vector of census counts most likely.

```{r}
sum(log(dbinom(x = dat_all$WIWA, size = 6, prob = 0.024)))


```

**Q6 (1 pt.): What are the two parameters for a binomial distribution and what do they represent?**

*Answer: The sample size and the probability*

**Q7 (2 pts.): What were the parameter values you selected?**

*Answer: n = , p= 0.24*

**Q8 (1 pt.): How did you choose a value for n?**

*Answer: *

**Q9 (2 pts.): Show R code you used to calculate the binomial log-likelihood for the vector of Winter Wren census counts.**

*Answer: *
```{r}
sum(log(dbinom(x = dat_all$WIWA, size = max(dat_bird$WIWR), prob = 0.24)))
```

**Q10 (1 pt.): Considering a Maximum Likelihood criterion, which model better fit the data?**



**Q11 (2 pts.): Considering what you know about the Binomial and Poisson distributions, which model is more appropriate for census count data?**


```{r}
set.seed(1)
vec_rnorm = rnorm(n = 10, mean = 0, sd = 1)
vec_rnorm


log(dnorm(vec_rnorm, 0, 1))
```


**Q12 (1 pt.): Create the vector (make sure you set seed to 1) and calculate the log likelihood that it came from a standard normal distribution.**

*Answer: log likelihood for every value of the vector are: -1.1151607, -0.9358010, -1.2680761, -2.1913990, -0.9732262, -1.2555227, -1.0377321, -1.1915002, -1.0847006, -0.9655696*

```{r}
set.seed(1)
vec_rnorm = rnorm(n = 10, mean = 0, sd = 1)
vec_rnorm


log(dnorm(vec_rnorm, 0, 1))

sum(log(dnorm(vec_rnorm, 0, 1)))
```


**Q13 (2 pts.): Show the R code you used to make the calculation**

```{r}
set.seed(1)
vec_rnorm = rnorm(n = 10, mean = 0, sd = 1)
vec_rnorm


log(dnorm(vec_rnorm, 0, 1))

sum(log(dnorm(vec_rnorm, 0, 1)))
```


**Q14 (2 pts.): Use the guess and check method to find the maximum likelihood optimal values for mean and standard deviation (to 2 significant figures) of the numbers in the vector.**

```{r}
set.seed(1)
vec_rnorm = rnorm(n = 10, mean = 0, sd = 1)

sum(log(dnorm(vec_rnorm, 0.1, 1)))

```

Load the Palmer penguins dataset.

Assume the data for flipper length are normally-distributed (they are not).

Determine maximum likelihood estimates for mean and standard deviation of a normal distribution that best describes the flipper lengths.

```{r}
library(palmerpenguins)
data("penguins")

mean(penguins$flipper_length_mm, na.rm = TRUE)
sd(penguins$flipper_length_mm, na.rm = TRUE)

sum(log(dnorm(!is.na(penguins$flipper_length_mm), mean = 200, sd= 14.06)))
```


**Q15 (1 pt.): How did you choose your starting guess values for mean and standard deviation?**



**Q16 (2 pts.): Using the guess and check method, what were the maximum likelihood estimates of mean and standard deviation (to 1 significant figure).**

```{r}
sum(log(dnorm(!is.na(penguins$flipper_length_mm), mean = 1, sd= 1)))
```



