---
title: "In-Class Models 1"
author: "David Murillo"
date: "November 1, 2022"
output:
  html_document:
    theme: cosmo
  html_notebook: default
---

```{r}
require(palmerpenguins)
dat_ade = droplevels(subset(penguins, species == "Adelie"))

hist(dat_ade$body_mass_g, main = "Adelie Penguins: Body Mass", xlab = "body mass (g)")

boxplot(body_mass_g ~ sex, data= penguins)
```

## Questions

# **Q1 (2 pts.): Create separate boxplots of body mass for male and female Adelie penguins. Your boxplots do not have to be in the same panel like mine. Show the R-code you used to make the plots.**

```{r}
Adelie <- droplevels(subset(penguins, species == "Adelie"))

Adelie_female <- droplevels(subset(penguins, sex == "female"))

Adelie_male <- droplevels(subset(penguins, sex == "male"))

par(mfrow = c(1,2))

boxplot(Adelie_female$body_mass_g, main = "Body mass of Adelie female",
        ylab = "Body mass", xlab= "")

boxplot(Adelie_male$body_mass_g, main = "Body mass of Adelie male",
        ylab = "Body mass", xlab= "")

```


- **Q2 (2 pts.): Perform a one-sample t-test of the alternative hypothesis that female Adelie penguins have a body mass different from zero grams. Note that this is a very silly alternative hypothesis. Is this a one- or two-tailed test? Show your R-code.**

*Answer: Is a two tailed test*

```{r}
Adelie_female <- droplevels(subset(penguins, sex == "female"))
t.test(Adelie_female$body_mass_g, alternative = "two.sided", mu= 0 )
```

- **Q4 (2 pts.): Now, conduct a slightly less silly test: perform a one-sample t-test of the alternative hypothesis that male Adelie penguins have a mean body mass greater than 4000 grams. Is this a one- or two-tailed test?**

*Answer: Is one tailed test*

```{r}
Adelie_male <- droplevels(subset(penguins, sex == "male"))
t.test(Adelie_male$body_mass_g, alternative = "greater", mu = 4000)
```

- **Q5 (1 pt.): Describe your conclusions based on the p-value of the t-test.**

- **Q6 (2 pts.): Conduct a two-sample t-test of the alternative hypothesis that male and female Adelie penguins have different mean body masses. Show your r-code.**

*Answer: *

```{r}
t.test(Adelie_male$body_mass_g, Adelie_female$body_mass_g)
```

- **Q7 (1 pt.): Describe your conclusions based on the p-value of the t-test.**

- **Q8 (2 pts.): Conduct a two-sample (one-tailed) t-test of the directional alternative hypothesis that male Adelie penguins are heavier than females.**

```{r}

t.test(Adelie_male$body_mass_g, Adelie_female$body_mass_g, alternative = "greater")

```

- **Q9 (2 pts.): Conduct a two-sample (one-tailed) t-test of the directional alternative hypothesis that male Adelie penguins are lighter than females.**

```{r}
t.test(Adelie_male$body_mass_g, Adelie_female$body_mass_g, alternative = "less")
```



