---
title: "David's Project Homepage"
author: "David Murillo"
date: '2022-12-02'
output: html_document
---

```{r setup, include=FALSE}
options(warn=-1)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warn = FALSE)
options(knitr.duplicate.label = "allow")
source("links.R")
```

```{r, include=FALSE}
# Load packagues

library(tidyverse)
library(lubridate)
library(sf)
library(raster)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(rio)
library(rgeos)
library(here)
```

```{r, include=FALSE}
# Load dataset

momotus <- read_csv2(here("data", "HnMomots.csv"))
```


## Welcome to David's Project Homepage

In this website we will see some birds distribution in Honduras.

# Honduras birds distribution {.tabset .tabset-pills}

## Momotidae

We will see a Momotidae family distribution in Honduras. 


The family Momotidae is endemic from the Neotropics region, and in Honduras there are 6 species of momoms. Momotus lessoni is the same that Momotus momota.

We will load a layer from Honduras

```{r}
Honduras <- ne_countries("medium", returnclass = "sf", country = "Honduras")
```

```{r}
ggplot(data= Honduras)+
  geom_sf()+
  geom_point(data = momotus, aes(decimalLongitude, decimalLatitude, color = species))+
  theme_bw()
```


## Golden Wing Warble (*Vermivora chrysoptera*)

```{r, echo= FALSE, warning=FALSE}

VERCHR <- read_csv(here("data", "VERCHR.csv"))
HN_VER <- filter(VERCHR, countryCode =="HN")



North_America <- ne_countries("small", returnclass = "sf", continent = "North America")

Honduras <- getData("GADM", country = "Honduras", level = 1) %>%
  st_as_sf()


Central_America <- ne_countries("medium", returnclass = "sf", 
                                country  = c("Honduras", "Nicaragua"))

ggplot()+
  geom_sf(data = Honduras) +
  geom_point(data = HN_VER, aes(Longitude, Latitude, color = individualCount)) +
  annotation_north_arrow(location = "tr", width = unit(1,"cm" ), 
                                          height = unit(1, "cm")) +
  annotation_scale(location= "br")+
  labs(title = "Distribution",
       subtitle = "Vermivora chrysopter",
       color= "Abundance")+
  theme(plot.title = element_text(face= "italic"))



```


